Rapid and Accurate Global PTM Discovery (G-PTM-D) Using Post-Acquisition Spectral Calibration and Multi-Notch Searches

### 2017-07-10 paper.tex
fjdkfj

A critical parameter for peptide identification is mass accuracy\cite{Scherl_2008}.
Higher mass accuracy provides increased specificity and thus confidence in peptide identifications, decreasing the false discovery rate.
Instrument noise, systemic drift and miscalibration limit the mass accuracy in acquired spectra.
Multiple calibration strategies to improve mass accuracy have been devised, and fall into three general categories:  External calibration prior to the MS experiment (e.g. standard instrument calibration protocols); internal calibration during the MS experiment (e.g. real-time calibration using a lock mass standard\cite{Olsen_2005}); and subsequent to the MS experiment (post-acquisition spectral calibration).
We use a post-acquisition calibration procedure that builds upon the software lock mass concept\cite{Cox_2011} recently reported by the Mann group.
In our strategy, the m/z differences between expected and observed peaks in the peptide tandem ms spectra are compiled, and then used to recalibrate the spectra.
%The increased mass accuracy of the calibrated spectra leads directly to improved identifications of both modified and unmodified peptides, as well as to increased confidence for PTM localization.

The spectral calibration procedure consists of two steps: Step 1: A narrow tolerance (e.g. 10 ppm precursor, 0.01 Dalton fragment) database search is performed on the dataset.
This yields a set of peptide identifications subject to a desired false discovery rate (e.g 1\%; based upon the target:decoy strategy \cite{Elias_2007}).
Step 2: The identifications from Step 1 are used to extract peak matches from the spectra, and a peak shifting procedure is performed.
The calibration algorithm is described in detail in Supplementary Material 1.

Calibration quality is accessed through the results of a standard 10ppm precursor tolerance database search.
PSMs within 1\% FDR prior to calibration are compared with PSMs after the calibration procedure, see Table~\ref{tbl:calib}.
The calibration procedure successfully addresses instrument miscalibration by centering the average error around zero.
The systemic noise introduced by various other factors is also addressed: this is evident in the decrease in the standard deviation of the errors in the precursor mass estimates, which places calibrated data squarely in the sub-ppm mode.
Histograms of the mass errors before and after calibration are plotted in Figure~\ref{fgr:figure1}.

>>>>

A critical parameter for peptide identification is mass accuracy\cite{Scherl_2008}.
Higher mass accuracy provides increased specificity and thus confidence in peptide identifications, decreasing the false discovery rate.
Instrument noise, systemic drift and miscalibration limit the mass accuracy in acquired spectra.
Multiple calibration strategies to improve mass accuracy have been devised, and fall into three general categories: External calibration prior to the MS experiment (e.g. standard instrument calibration protocols); internal calibration during the MS experiment (e.g. real-time calibration using a lock mass standard\cite{Olsen_2005}); and spectral calibration subsequent to the MS experiment (post-acquisition spectral calibration).
We developed a post-acquisition calibration procedure that builds upon the software lock mass concept\cite{Cox_2011} recently reported by the Mann group.
In our strategy, the m/z differences between expected and observed peaks in the peptide tandem ms spectra are compiled, and then used to recalibrate the spectra.
The spectral calibration procedure consists of two steps: Step 1: A narrow tolerance (e.g. 10 ppm precursor, 0.01 Dalton fragment) database search is performed on the dataset.
This yields a set of peptide identifications subject to a desired false discovery rate (e.g 1\%; based upon the target:decoy strategy\cite{Elias_2007}).
Step 2: The identifications from Step 1 are used to extract peak matches from the spectra, and a peak shifting procedure is performed.
The calibration algorithm is described in detail in Supplementary Material 1.

Calibration quality is evaluated through the results of a standard 10ppm precursor tolerance database search.
PSMs within 1\% FDR prior to calibration are compared with PSMs after the calibration procedure (Table~\ref{tbl:calib}).
The calibration procedure successfully addresses instrument miscalibration by centering the average error around zero.
The systemic noise introduced by various other factors is also addressed: this is evident in the decrease in the standard deviation of the errors in the precursor mass estimates, which show that the calibrated mass spectra have sub-ppm mass errors.
Histograms of the mass errors before and after calibration are plotted in Figure~\ref{fgr:figure1}.
